---
title: "Lab05"
author: "Haoran Zhang"
date: "2021/9/24"
output: 
  html_document: 
    keep_md: yes
---

```{r}
library(data.table)
library(dplyr)
library(dtplyr)
```

## Setup
dat
```{r}
# Download the data
# Where are we getting the data from
met_url <- "https://github.com/USCbiostats/data-science-data/raw/master/02_met/met_all.gz"
# Downloading the data to a tempfile (so it is destroyed afterwards)
# you can replace this with, for example, your own data:
# tmp <- tempfile(fileext = ".gz")
tmp <- "met.gz"
# We sould be downloading this, ONLY IF this was not downloaded already.
# otherwise is just a waste of time.
if (!file.exists(tmp)) {
  download.file(
    url      = met_url,
    destfile = tmp,
    # method   = "libcurl", timeout = 1000 (you may need this option)
  )
}
dat <- fread(tmp)
head(dat)
```
stations
```{r}
stations <- fread("isd-history.csv")
stations[, USAF := as.integer(USAF)]

# Dealing with NAs and 999999
stations[, USAF   := fifelse(USAF == 999999, NA_integer_, USAF)]
stations[, CTRY   := fifelse(CTRY == "", NA_character_, CTRY)]
stations[, STATE  := fifelse(STATE == "", NA_character_, STATE)]

# Selecting the three relevant columns, and keeping unique records
stations <- unique(stations[, list(USAF, CTRY, STATE)])

# Dropping NAs
stations <- stations[!is.na(USAF)]

# Removing duplicates
stations[, n := 1:.N, by = .(USAF)]
stations <- stations[n == 1,][, n := NULL]

```
Merging the data
```{r}
met <- merge(
  # Data
  x     = dat,      
  y     = stations, 
  # List of variables to match
  by.x  = "USAFID",
  by.y  = "USAF", 
  # Which obs to keep?
  all.x = TRUE,      
  all.y = FALSE
  )
```


## Question 1: Representative station for the US
```{r}
station_averages<-met[,.(
  temp=mean(temp,na.rm = TRUE),
  wind.sp=mean(wind.sp,na.rm = TRUE),
  atm.press=mean(atm.press,na.rm = TRUE),
  lat=mean(lat,na.rm = TRUE),
  lon=mean(lon,na.rm = TRUE)
),by="USAFID"]
```

```{r}
medians<-station_averages[,.(
  temp_50=quantile(temp,probs = 0.5,na.rm = TRUE),
  wind.sp_50=quantile(wind.sp,probs = 0.5,na.rm = TRUE),
  atm.press_50=quantile(atm.press,probs = 0.5,na.rm = TRUE)
)]
medians
```
Find stations that are closest. (which.min())
```{r median-temp}
station_averages[,temp_dist:=abs(temp-medians$temp_50)]
median_temp_station<-station_averages[order(temp_dist)][1]
```
The median temperature station is `r median_temp_station$USAFID`.

## Question 2: Representative station per state
First recover state variable by merging
```{r}
station_averages<-merge(x = station_averages, y = stations,by.x = "USAFID", by.y = "USAF", all.x = TRUE, all.y = FALSE)
station_averages
station_averages[, temp_50:=quantile(temp,probs = 0.5,na.rm = TRUE), by=STATE]
station_averages[, wind.sp_50:=quantile(wind.sp,probs = 0.5,na.rm = TRUE), by=STATE]
station_averages[, atm.press_50:=quantile(atm.press,probs = 0.5,na.rm = TRUE), by=STATE]

station_averages
```
Euclidean $\sqrt{\sum_i(x_i-y_i)^2}$
```{r}
station_averages[,eudist:=sqrt(
  (temp-temp_50)^2+(wind.sp-wind.sp_50)^2
)]
```

Find representitive for each state
```{r}
rep_state <- station_averages[ , .SD[which.min(eudist)], by = STATE]
rep_state
```
## Question 3: In the middle?
```{r}
midpoint <- met[, .(
  lon_50 = quantile(lon, probs = 0.5, na.rm = TRUE),
  lat_50 = quantile(lat, probs = 0.5, na.rm = TRUE)
), by = STATE]

station_averages <- merge(
  x = station_averages,
  y = midpoint,
  by = "STATE"
)
```

```{r}
station_averages[, mid_eudist:= sqrt(
  (lat - lat_50)^2+(lon - lon_50)^2
)]
station_averages
```

```{r}
mid_state <- station_averages[ , .SD[which.min(mid_eudist)], by = STATE]
mid_state
```
merge data of median and mid stations
```{r}
rep_state[, type := "median station"]
mid_state[, type := "mid point"]
all_stations <- rbind(rep_state, mid_state, fill = TRUE)
```

draw the map
```{r}
library(leaflet)
pal <- colorFactor(c("red","blue"), domain = all_stations$type)
leaflet() %>% 
  addProviderTiles('CartoDB.Positron') %>% 
  addCircles(
    data = all_stations,
    lat = ~lat, lng = ~lon, 
    opacity = 1, fillOpacity = 1, radius = 400, color = ~pal(all_stations$type)
    ) %>%
  addLegend('bottomleft', pal=pal, values=all_stations$type,
          title='Sites', opacity=1)
```

## Question 4: Means of means
```{r}
met[,state_temp:=mean(temp,na.rm = TRUE), by=STATE]
met[, state_wind.sp := mean(wind.sp, na.rm = TRUE),by = STATE]
met[, state_atm.press := mean(atm.press, na.rm = TRUE),by = STATE]
met[,temp_cat:=fifelse(state_temp<20,"low-temp",
                       fifelse(state_temp<25,"mid-temp",
                               "high-temp"))]
table(met$temp_cat,useNA = "always")
```
summarize
```{r}
tab <- met[, .(
  N_entries = .N,
  N_stations = length(unique(USAFID)),
  avg_temp = mean(state_temp),
  avg_wind.sp = mean(state_wind.sp),
  avg_atm.press = mean(state_atm.press)
), by = temp_cat]

knitr::kable(tab)
```





